\chapter*{Abstract}

The relatively recently proposed hierarchical phrase-based
translation model for statistical machine translation (SMT)
has achieved state-of-the-art performance
in numerous recent translation evaluations. Hierarchical phrase-based
systems comprise a pipeline of modules with complex interactions.
In this thesis, we propose refinements to the hierarchical
phrase-based model as well as improvements and analyses in various
modules for hierarchical phrase-based systems.

We took the opportunity of increasing amounts of available training
data for machine translation as well as existing framework for
distributed computing in order to build better infrastructure
for extraction, estimation and retrieval of hierarchical phrase-based
grammars.

The modular nature of the SMT pipeline, while allowing individual
improvements, has the disadvantage that errors committed by
one module are propagated to the next. This thesis
alleviates this issue between the word alignment module and the
grammar extraction and estimation module by considering richer
statistics from word alignment models in extraction.

This thesis also proposes refinements in grammar and language
modelling both in the context of domain adaptation and in the
context of the interaction between first-pass decoding and
lattice rescoring.

Finally, the fluency of machine translation output is studied
through the string regeneration task. State-of-the-art is achieved
in the string regeneration task. String regeneration applied
to the output of a hierarchical phrase-based translation system
obtains the same level of translation quality.
