\chapter*{Acknowledgements}

First, I am truly indebted to my supervisor
Prof.\ Bill Byrne.
Bill provided relentless support throughout
my PhD studies.
He taught me how to be a better writer and a better researcher.
I am also very grateful that he provided me with
invaluable opportunities such as participating
in translation competitions or carrying out an internship
at Google for three months.

I also would like to thank Dr.\ Stephen Clark and Dr.\ Miles
Osborne for being part of my thesis committee and providing
very valuable feedback on the first submission of this thesis.

Without the lab Computer Officers, Patrick Gosling
and Anna Langley, experiments reported in this thesis
would not have been possible, I am therefore
grateful for all their hard work on the computer
systems.

I also wish to thank all my colleagues at
the Machine Intelligence Lab.
Dr.\ Adrià de Gispert provided me with
critical guidance throughout my research: two chapters
of this thesis originated from a very fruitful collaboration
with him.
I am grateful to Aurelien Waite for all his
help on infrastructure and for helping me becoming
a better engineer.
Thanks to Dr.\ Gonzalo Iglesias for illuminating
discussions on FSTs for machine translation as well
as C++ software engineering.
Dr.\ Graeme Blackwood and Dr.\ Jamie Brunning
provided much support at the start of my studies as
well as critical translation tools such as word alignment
and lattice rescoring tools.

I am also in debt to my office colleagues Matt Shannon, Matt Seigel
and Chao Zhang for making PhD studies a pleasurable experience.

Finally, thank you Anna for your infinite patience, guidance and
support and for telling me to apply for at least two jobs at
the end of my studies. Thank you Sofia and Fania: I will
try to teach you many languages just in case machine translation
doesn't improve quickly enough.
